Fundamental Problems with the Evidence Base of Adolescent Depression Treatments.

### 2024-01-04 apples_and_oranges.qmd
several changes with brining functions up etc.

@@

# I have created a function for this
library(tidyverse)
library(truncnorm)

@@

simulate_dataframes_for_st_errors <- function(df, num_repetitions, seed, n, a, b, mean, sd) { # n refers to the number
                                                                                              # of unique ids to which a correlation coef
                                                                                              # is allocated.
  set.seed(seed)

@@

@@

# create a function for metaregression

run_metaregression <- function(list_of_datasets, model_formula) {
  list_model_meta_reg <- list()

model_1 <- as.formula(~ four_level_var)
list_model_1_meta_reg <- run_metaregression(list_df_simulated, model_1)

# list_model_1_meta_reg[[2]]

# Function to extract parameters (coefs, ses, CI) for each level of the dummy variable
# it is an awkward one, because I couldnt' come up with a better way to add the intercept to each coefficient
# given the named output of the built in coef function. 
# here I exctract the coeffcients
extract_coefficients_func <- function(df_with_coefs){
  list_coefs <- list()
  coefs <- 0
  coefficient_output <- 0

# dummy_var_means <- extract_coefficients_func(list_model_1_meta_reg[[1]])

list_dummy_var_means <-lapply(list_model_1_meta_reg, extract_coefficients_func)

# this is the way to get the means and SEs from the simulation across all datasets.

df_mean_coefs_from_sim <- data.frame(

  condition = levels(list_df_simulated[[1]]$four_level_var),

  mean_coefficient = c(mean(sapply(list_dummy_var_means, function(df) df[1,1])),
                      mean(sapply(list_dummy_var_means, function(df) df[2,1])),
                      mean(sapply(list_dummy_var_means, function(df) df[3,1])),
                      mean(sapply(list_dummy_var_means, function(df) df[4,1]))),

  mean_se = c(mean(sapply(list_dummy_var_means, function(df) df[1,2])),
              mean(sapply(list_dummy_var_means, function(df) df[2,2])),
              mean(sapply(list_dummy_var_means, function(df) df[3,2])),
              mean(sapply(list_dummy_var_means, function(df) df[4,2])))

)

# and add the CIs

df_mean_coefs_from_sim$upper_ci <- df_mean_coefs_from_sim$mean_coefficient + 1.96 * df_mean_coefs_from_sim$mean_se

df_mean_coefs_from_sim$lower_ci <- df_mean_coefs_from_sim$mean_coefficient - 1.96 * df_mean_coefs_from_sim$mean_se
# df_mean_coefs_from_sim # this is now the dataframe containing the means you need for plotting.

#also, add the n of studies non missing from variable created above
#important first create the same condition variable: 
df_count_studies_not_missing$condition <- str_replace_all(df_count_studies_not_missing$condition, "_", " ")

# I needed to join them because of the relevelling, the rows would not align otherwise.
# df_mean_coefs_from_sim <- right_join(df_mean_coefs_from_sim, df_count_studies_not_missing)

# need to merge this with the count studies variable to obtain the ks
library(forcats)
ordering_criterion <- unique(df_mean_coefs_from_sim$condition)

# 
df_count_studies_not_missing <- df_count_studies_not_missing %>% arrange(ordering_criterion) %>%
  mutate(condition = fct_relevel(condition, levels = unique(condition)))
# I will use this for plotting below.

@@

NOTE: **Charlotte**, this should be moved to the supplement at Table 2
```{r, warning=FALSE, message = FALSE, echo=FALSE}
variability_in_sims <- data.frame (mean_coefficient = c(mean(sapply(list_dummy_var_means, function(df) df[1,1])),
                      mean(sapply(list_dummy_var_means, function(df) df[2,1])),
                      mean(sapply(list_dummy_var_means, function(df) df[3,1])),
                      mean(sapply(list_dummy_var_means, function(df) df[4,1]))),

sd_coefficient = c(sd(sapply(list_dummy_var_means, function(df) df[1,1])),
                      sd(sapply(list_dummy_var_means, function(df) df[2,1])),
                      sd(sapply(list_dummy_var_means, function(df) df[3,1])),
                      sd(sapply(list_dummy_var_means, function(df) df[4,1]))),

  mean_se = c(mean(sapply(list_dummy_var_means, function(df) df[1,2])),
              mean(sapply(list_dummy_var_means, function(df) df[2,2])),
              mean(sapply(list_dummy_var_means, function(df) df[3,2])),
              mean(sapply(list_dummy_var_means, function(df) df[4,2]))),

  sd_se = c(sd(sapply(list_dummy_var_means, function(df) df[1,2])),
              sd(sapply(list_dummy_var_means, function(df) df[2,2])),
              sd(sapply(list_dummy_var_means, function(df) df[3,2])),
              sd(sapply(list_dummy_var_means, function(df) df[4,2])))
)

@@

# Create summary stats text
df_mean_coefs_from_sim$text_label <- 0
for(i in 1: nrow(df_mean_coefs_from_sim))
df_mean_coefs_from_sim$text_label[i] <-  paste(
   "k = ", df_count_studies_not_missing $n[i],"\n", 
   round(df_mean_coefs_from_sim$mean_coefficient[i], 2),
   "[" ,
 round(df_mean_coefs_from_sim$lower_ci[i],2), 
 round(df_mean_coefs_from_sim$upper_ci[i], 2),
"]")

@@

plot_means <- ggplot(df_mean_coefs_from_sim, aes(x = mean_coefficient, y = 1:nrow(df_mean_coefs_from_sim),
                                   colour = condition, label = text_label)) +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = lower_ci, xmax = upper_ci), width = 0.2, position = position_dodge(0.5)) +
  geom_text(vjust = +1.5, size = 4) +  # Adjust vjust and size as needed
  scale_size_continuous(guide = "none") +
  guides(colour = FALSE) + 
  scale_color_manual(values = c(redish_palette, blueish_palette)) +
  theme_minimal() +
  labs(x = "TE-random", y = NULL, title = "Adolescent Depression Trial Efficacy by Treatment Arm",
       subtitle = "metanalytically derived estimates of within group changes") +
  xlab("Standardized Mean Difference (SMD) with 95% CIs") +
  ylab("") +
  ylim(0, nrow(df_mean_coefs_from_sim) + 1) +
  xlim(-2.5, 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5, colour = "grey") +
  geom_segment(x = -1.25, xend = -1.75, y = 4.7, yend = 4.7, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = -1.55, y = 4.9, label = "More Effective", color = "grey", vjust = 0.5, hjust = 1) +
  geom_segment(x = -1.35, xend = -0.85, y = 4.9, yend = 4.9, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = -1.05, y = 5.05, label = "Less Effective", color = "grey", vjust = 0.5, hjust = 0)+
  theme(axis.title.x = element_text(size = 14), 
        axis.text.x  = element_text(size = 14),
        axis.text.y = (element_blank()),
        plot.title = element_text(size = 16),
        plot.subtitle = element_text(size = 14))+
  geom_curve(aes(x = 0.18, y = 4.73, xend = 0.02, yend = 4.5), color = "grey", curvature = -0.2, arrow = arrow(length = unit(0.25, "cm"), type = "closed")) +
  geom_text(x = 0.2, y = 4.75, label = "Line of No Effect", color = "grey", vjust = 0.5, hjust = 0)

>>>>

```{r, warning=FALSE, message = FALSE, echo=FALSE}
###########Libraries needed
library(tidyverse)
library(truncnorm)
library(metafor)
```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#############Some custom made functions#############

###### 1. Simulation Function for SEs
simulate_dataframes_for_st_errors <- function(df, num_repetitions, seed, n, a, b, mean, sd) { # n refers to the number
                                                                                              # of unique ids to which a correlation coef
                                                                                              # is allocated.
  set.seed(seed)

###############2. Simulation function for metaregression
run_metaregression <- function(list_of_datasets, model_formula) {
  list_model_meta_reg <- list()

###########3. Function to extract parameters (coefs, ses, CI) for each level of the dummy #variable
# it is a slightly awkward one, because I couldnt' come up with a better way to add the intercept to #each coefficient
# given the named output of the built in coef function. 
# here I exctract the coeffcients
extract_coefficients_func <- function(df_with_coefs){
  list_coefs <- list()
  coefs <- 0
  coefficient_output <- 0

############4. Function to get the means out of the extracted parameters
# here the argument list_of_dfs is the list of extracted coefficients that you get out of the #extract_coefficients_func
# which itself you get from the run_metaregression function. The argument conditions refers to #the levels of the variable that describes each study arm

calculate_mean_coefs_ses <- function(list_of_dfs, conditions) {
  df_coefs <- matrix(NA, nrow = length(list_of_dfs), ncol = length(conditions))
  df_se <- matrix(NA, nrow = length(list_of_dfs), ncol = length(conditions))

  for (i in 1:length(list_of_dfs)) {
    for (j in 1:length(conditions)) {
      df_coefs[i, j] <- list_of_dfs[[i]]$coefficients[j]
      df_se[i, j] <- list_of_dfs[[i]]$se[j]
    }
  }

  df_mean_coefs_ses <- data.frame(
    coef_means = colMeans(df_coefs),
    se_means = colMeans(df_se)
  )

  df_mean_coefs_ses$lower_ci <- df_mean_coefs_ses$coef_means - 1.96 * df_mean_coefs_ses$se_means
  df_mean_coefs_ses$upper_ci <- df_mean_coefs_ses$coef_means + 1.96 * df_mean_coefs_ses$se_means

  return(df_mean_coefs_ses)
}

@@

# Use the function I have created for this

@@

@@

  df_count_studies$condition <- factor(df_count_studies$condition) # turn to factor

  # relevel so that medication control becomes the reference category for the regression
  df_count_studies$condition <- relevel(df_count_studies$condition, ref = "medication_control")

@@

######## NOW RUN METAREGRESSIONS AND GET Coefficients for 
# A) the overall sample
# B) the CBT vs Fluox and Escitalopram sample
# C) the no WL sample

# A. Estimate Means and CIs for the overall ----------------------------------
# specify model
model_1 <- as.formula(~ four_level_var)

# run metareg function 
# here you can use the simulated results directly
list_model_1_meta_reg <- run_metaregression(list_df_simulated, model_1)

# extract coefs from the  list
list_dummy_var_means <-  lapply(list_model_1_meta_reg, extract_coefficients_func)

# calculate mean coefs and ses
df_mean_coefs_from_sim <- calculate_mean_coefs_ses(means_cbt_fluox_esc_study, conditions)

 df_mean_coefs_from_sim <- data.frame(cbind(condition,df_mean_coefs_from_sim))

# B. Estimate Means and CIs for CBT Fluox Esc --------------------------------
# first grab ids for those studies
cbt_fluox_esc_study_ids <-
  df_appl_v_orange[df_appl_v_orange$active_type == "cbt" | df_appl_v_orange$active_type == "Fluoxetine" | df_appl_v_orange$active_type == "Escitalopram",]$new_study_id

# use those ids to subset all dataframes in the list
list_cbt_fluox_esc_study <- list()

model_1 <- as.formula(~ four_level_var)

# apply the metareg function to the list of cbt, fluox, esc
list_model_1_cbt_fluox_esc_study <- run_metaregression(list_cbt_fluox_esc_study , model_1)

# extract coefs from the list for cbt fluox escit
means_cbt_fluox_esc_study <-  lapply(list_model_1_cbt_fluox_esc_study, extract_coefficients_func)

# calculate mean coefs and ses
conditions <- unique(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_cbt_fluox_esc_study <- calculate_mean_coefs_ses(means_cbt_fluox_esc_study, conditions)
 coef_and_se_means_cbt_fluox_esc_study <- data.frame(cbind(condition, coef_and_se_means_cbt_fluox_esc_study))

# C. Estimate Means and CIs without WL ---------------------------------------
# grab the ids for no waitlist
no_wl_ids <-
  df_appl_v_orange[df_appl_v_orange$control_type != "wl",]$new_study_id

# use them to loop over the simulated list to subset all the dataframes.
list_no_wl <- list()

  list_no_wl[[i]] <- list_df_simulated[[i]][list_df_simulated[[i]]$new_study_id
                                            %in% no_wl_ids,]

# apply the function to the list of cbt, fluox, esc
list_model_1_no_wl <- run_metaregression(list_no_wl , model_1)

# now use the function to extract the coefficients
means_no_wl <-  lapply(list_model_1_no_wl, extract_coefficients_func)

# run the mean function for no wl
conditions <- unique(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_no_wl <- calculate_mean_coefs_ses(means_no_wl, conditions)
coef_and_se_means_no_wl <- data.frame(cbind(condition,coef_and_se_means_no_wl))

 #add the ns of studies FIX THIS

 df_mean_coefs_from_sim <- data.frame(cbind(df_mean_coefs_from_sim, n = df_count_studies[df_count_studies$is_missing == FALSE, ]$n))

# here are all three of them OVERALL, CBT and WL together. We may want to display them together. 
# df_all_coefficients <- cbind(df_mean_coefs_from_sim, coef_and_se_means_cbt_fluox_esc_study,coef_and_se_means_no_wl)

@@

@@

# Create summary stats text
df_mean_coefs_from_sim$text_label <- 0
for(i in 1: nrow(df_mean_coefs_from_sim))
df_mean_coefs_from_sim$text_label[i] <-  paste(
   "k = ", df_mean_coefs_from_sim$n[i],"\n", 
   round(df_mean_coefs_from_sim$coef_means[i], 2),
   "[" ,
 round(df_mean_coefs_from_sim$lower_ci[i],2), 
 round(df_mean_coefs_from_sim$upper_ci[i], 2),
"]")

@@

plot_means <- ggplot(df_mean_coefs_from_sim, aes(x = coef_means, y = 1:nrow(df_mean_coefs_from_sim),
                                   colour = condition, label = text_label)) +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = lower_ci, xmax = upper_ci), width = 0.2, position = position_dodge(0.5)) +
  geom_text(vjust = +1.5, size = 4) +  # Adjust vjust and size as needed
  scale_size_continuous(guide = "none") +
  guides(colour = FALSE) + 
  scale_color_manual(values = c(redish_palette, blueish_palette)) +
  theme_minimal() +
  labs(x = "TE-random", y = NULL, title = "Adolescent Depression Trial Efficacy by Treatment Arm",
       subtitle = "metanalytically derived estimates of within group changes") +
  xlab("Standardized Mean Difference (SMD) with 95% CIs") +
  ylab("") +
  ylim(0, nrow(df_mean_coefs_from_sim) + 1) +
  xlim(-2.5, 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5, colour = "grey") +
  geom_segment(x = -1.25, xend = -1.75, y = 4.7, yend = 4.7, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = -1.55, y = 4.9, label = "More Effective", color = "grey", vjust = 0.5, hjust = 1) +
  geom_segment(x = -1.35, xend = -0.85, y = 4.9, yend = 4.9, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = -1.05, y = 5.05, label = "Less Effective", color = "grey", vjust = 0.5, hjust = 0)+
  theme(axis.title.x = element_text(size = 14), 
        axis.text.x  = element_text(size = 14),
        axis.text.y = (element_blank()),
        plot.title = element_text(size = 16),
        plot.subtitle = element_text(size = 14))+
  geom_curve(aes(x = 0.18, y = 4.73, xend = 0.02, yend = 4.5), color = "grey", curvature = -0.2, arrow = arrow(length = unit(0.25, "cm"), type = "closed")) +
  geom_text(x = 0.2, y = 4.75, label = "Line of No Effect", color = "grey", vjust = 0.5, hjust = 0)
